---
title: "STARS_SSN_join_test"
author: "Luke Vawter"
date: '2023-02-28'
output: html_document
---

testing if we can extract data, link it to ids and put it back

Andrea:
"Both the 1_Preds and 4_Preds folders contain 2 subfolders: ssn and data. Hopefully the data can be joined to the ssn on STATIONID in R. 

1_Preds has all data at the predication sites in one file, whereas 
4_Preds has four preds files.

See Data_descriptions for an explanation of the abbreviations used in the data files."

import SSN package
```{r}
library("SSN")

```

```{r}
path_1preds = "data/push_test/Separate_Data/1Preds/lsn1/lsn.ssn"
# path_4preds = "data/push_test/Separate_Data/4Preds/lsn/lsn.ssn"

path_1data_preds = "data/push_test/Separate_Data/1Preds/data1/preds_data/preds_data.txt"

path_1data_sites = "data/push_test/Separate_Data/1Preds/data1/sites_data.txt"

# path_4data_sites = "data/push_test/Separate_Data/4Preds/data1//sites_data.txt"


```

Import the data into an SSN object and pull out to dataframe
```{r}
ssn_1preds.ssn <- importSSN(path_1preds)#, predpts = "preds")

ssn_1preds.df<- getSSNdata.frame(ssn_1preds.ssn)
```

get site data to dataframe
```{r}
data1_sites.df <- data.table::fread(path_1data_sites,  data.table = FALSE)
```

join site and
```{r}
library(dplyr)
ssn_1preds_joined.df <- left_join(ssn_1preds.df, data1_sites.df, by = c("STATIONID"))# %>%
  # select(-"")
  
# ssn_1preds_joined_dif.df <- left_join(ssn_1preds.df, data1_sites.df, by = c("SITE","SDATE")) #%>%
  # select(-"pid")

unique_ssn_1preds_joined.df <- unique(ssn_1preds_joined.df)
```

```{r}
ssn_1preds_joined.ssn <- putSSNdata.frame(ssn_1preds_joined.df, ssn_1preds.ssn)

ssn_1preds_joined2.ssn <- putSSNdata.frame(ssn_1preds.ssn, ssn_1preds_joined.df)

help(putSSNdata.frame)

```

```{r}
test <- (!"pid" %in% names(ssn_1preds_joined.df)| !all(row.names(ssn_1preds_joined.df) == ssn_1preds_joined.df$pid))
```

```{r}
test1 <- (!"pid" %in% names(ssn_1preds_joined.df))
test2 <- !all(row.names(ssn_1preds_joined.df) == ssn_1preds_joined.df$pid)
test3 <- ssn_1preds_joined.df$pid

row.names(ssn_1preds_joined.df) <-as.character(row.names(ssn_1preds_joined.df))
typeof(row.names(ssn_1preds_joined.df))

row.names(ssn_1preds_joined.df) <- ssn_1preds_joined.df$pid
```


